Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.Rmd
update the text to match the updated plot

The power of visual test arises steadily as $log_e(E)$ increases from 2 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features. This can also be observed from the example residual plots that when $log_e(E) = 2.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The visual pattern becomes much clearer as $log_e(E)$ increases. At $log_e(E) \approx 6$, the power reaches almost 100%.

The power of all inappropriate tests except for RESET test shows improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

@@

A primary factor contributes to the non-linearity model is the shape of the non-linearity. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have almost identical power in testing the "U", the "S" and the "M" shape, except for testing the "Triple-U" shape, which have significant lower power. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969). However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will have little difference to other shapes. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data structure in order to set the parameters for the test.

For visual tests, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "U" shape. This implies the readability of the shape do not strictly follow the degree of the polynomials.

@@

We also investigate the impact of different heteroskedasticity shapes on power of conventional tests and visual tests. In theory, the "Left-triangle" and the "Right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test. As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived. Similarly, visual tests will have the same power of detecting these two shapes if they are equally likely to be identified. However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies.

The chance of detecting the "Butterfly" shape is roughly the same as the chance of detecting the "Left-triangle" shape in both tests. The only noticeable difference from Figure \ref{fig:heter-power-uniform-a} is the Breusch-Pagan test has slightly higher power for detecting this shape compared to other two shapes when the effect size is low.

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model. We focus on $log(E) > 2$ where the visual patterns start to become recognisable.

>>>>

The power of visual test arises steadily as $log_e(E)$ increases from 3 to 5 for both non-linearity patterns and heteroskedasticity patterns, suggesting that the effect starts to make significant impact on the degree of the presence of the designed visual features. This can also be observed from the example residual plots that when $log_e(E) = 3.5$, a weak "S-shape" and a weak "triangle" shape are presented in Figure \ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The visual pattern becomes much clearer as $log_e(E)$ increases. At $log_e(E) = 5$, the power reaches almost 100%.

The power of all inappropriate tests show improvement as the effect increases but at a lower rate than the visual test in both scenarios. This coincides the point made by @cook1982residuals that residual-based tests for a specific type of model defect may be sensitive to other types of model defects. The power curve of RESET test remains at around 5% in Figure \ref{fig:heterpower} since there are no non-linear terms leave out in the heteroskedasticity model and $H_0$ of the test is always satisfied.

@@

A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have identical power in testing the "U", the "S" and the "M" shape. But it has significant lower power in testing the "Triple-U" shape. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969). However, the "Triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "Triple-U" shape will have little difference to other shapes. The recommendation of the polynomial power for the RESET should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we expect the "U" shape will be the easiest one to be detected by subjects followed by the "S", "M" and "Triple-U" shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power curves are mostly aligned with our expectation, except for the "M" shape, which is as easy to be detected as the "U" shape. This implies the readability of the shape do not strictly follow the degree of the polynomials.

@@

We have also investigated the impact of different heteroskedasticity shapes on power of conventional tests and visual tests. In theory, the "Left-triangle" and the "Right-triangle" shapes are functionally identical from the point of view of a Breusch-Pagan test. As shown in Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where little difference between the power curves can be perceived. Similarly, visual tests should have the smae power in detecting these two shapes if they are equally likely to be identified. However, it can be observed from Figure \ref{fig:heter-power-uniform-a} that the power curve for the "Left-triangle" shape is constantly higher than the one for the "Right-triangle" shape, indicating a potential favour of orientation by human, which is worth to be explored in future studies. Besides, the chance of detecting the "Butterfly" shape by visual tests are slightly higher than other two shapes.

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of power of visual tests and conventional tests for different fitted distributions used in the non-linearity model and the heteroskedasticity model. Note that we have collected more evaluations on lineup with uniform fitted value distribution, and the number of evaluations on a lineup will have an impact on the power of the visual test. To have a fair comparison, the five evaluations used in the power estimation are randomly sampled from the total eleven evaluations for each lineup with uniform fitted value distribution. The sampling process has been repeated 500 times to visualize the variation.

### 2023-02-07 paper_comparison/paper_comparison/paper_comparison.tex
update the text to match the updated plot

The power of visual test arises steadily as \(log_e(E)\) increases from
2 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 2.5\), a
weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
visual pattern becomes much clearer as \(log_e(E)\) increases. At
\(log_e(E) \approx 6\), the power reaches almost 100\%.

The power of all inappropriate tests except for RESET test shows
improvement as the effect increases but at a lower rate than the visual
test in both scenarios. This coincides the point made by
\citet{cook1982residuals} that residual-based tests for a specific type
of model defect may be sensitive to other types of model defects. The
power curve of RESET test remains at around 5\% in Figure
\ref{fig:heterpower} since there are no non-linear terms leave out in
the heteroskedasticity model and \(H_0\) of the test is always
satisfied.

@@

A primary factor contributes to the non-linearity model is the shape of
the non-linearity. According to Figure \ref{fig:poly-power-uniform-j},
conventional tests have almost identical power in testing the ``U'', the
``S'' and the ``M'' shape, except for testing the ``Triple-U'' shape,
which have significant lower power. To understand why this is, one needs
to return to the way the RESET test is applied. It requires a parameter
indicating degree of fitted values to test for, and the recommendation
is to generically use four (\citet{ramsey_tests_1969}). However, the
``Triple-U'' shape constructed from the Hermite polynomials use power up
to 18. If the RESET test had been applied using a higher power no less
than six, the power curve of ``Triple-U'' shape will have little
difference to other shapes. The recommendation of the polynomial power
for the RESET should be revised, perhaps. This illustrates the
sensitivity of conventional testing to the parameters, and it also
points to a limitation that one needs to know the data structure in
order to set the parameters for the test.

For visual tests, we expect the ``U'' shape will be the easiest one to
be detected by subjects followed by the ``S'', ``M'' and ``Triple-U''
shape. From Figure \ref{fig:poly-power-uniform-j}, it can be observed
that the power curves are mostly aligned with our expectation, except
for the ``M'' shape, which is as easy to be detected as the ``U'' shape.
This implies the readability of the shape do not strictly follow the
degree of the polynomials.

@@

We also investigate the impact of different heteroskedasticity shapes on
power of conventional tests and visual tests. In theory, the
``Left-triangle'' and the ``Right-triangle'' shapes are functionally
identical from the point of view of a Breusch-Pagan test. As shown in
Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
little difference between the power curves can be perceived. Similarly,
visual tests will have the same power of detecting these two shapes if
they are equally likely to be identified. However, it can be observed
from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
``Left-triangle'' shape is constantly higher than the one for the
``Right-triangle'' shape, indicating a potential favour of orientation
by human, which is worth to be explored in future studies.

The chance of detecting the ``Butterfly'' shape is roughly the same as
the chance of detecting the ``Left-triangle'' shape in both tests. The
only noticeable difference from Figure \ref{fig:heter-power-uniform-a}
is the Breusch-Pagan test has slightly higher power for detecting this
shape compared to other two shapes when the effect size is low.

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
power of visual tests and conventional tests for different fitted
distributions used in the non-linearity model and the heteroskedasticity
model. We focus on \(log(E) > 2\) where the visual patterns start to
become recognisable.

>>>>

The power of visual test arises steadily as \(log_e(E)\) increases from
3 to 5 for both non-linearity patterns and heteroskedasticity patterns,
suggesting that the effect starts to make significant impact on the
degree of the presence of the designed visual features. This can also be
observed from the example residual plots that when \(log_e(E) = 3.5\), a
weak ``S-shape'' and a weak ``triangle'' shape are presented in Figure
\ref{fig:polypower} and Figure \ref{fig:heterpower} respectively. The
visual pattern becomes much clearer as \(log_e(E)\) increases. At
\(log_e(E) = 5\), the power reaches almost 100\%.

The power of all inappropriate tests show improvement as the effect
increases but at a lower rate than the visual test in both scenarios.
This coincides the point made by \citet{cook1982residuals} that
residual-based tests for a specific type of model defect may be
sensitive to other types of model defects. The power curve of RESET test
remains at around 5\% in Figure \ref{fig:heterpower} since there are no
non-linear terms leave out in the heteroskedasticity model and \(H_0\)
of the test is always satisfied.

@@

A primary factor contributes to the non-linearity model is the shape of
the non-linearity pattern. According to Figure
\ref{fig:poly-power-uniform-j}, conventional tests have identical power
in testing the ``U'', the ``S'' and the ``M'' shape. But it has
significant lower power in testing the ``Triple-U'' shape. To understand
why this is, one needs to return to the way the RESET test is applied.
It requires a parameter indicating degree of fitted values to test for,
and the recommendation is to generically use four
(\citet{ramsey_tests_1969}). However, the ``Triple-U'' shape constructed
from the Hermite polynomials use power up to 18. If the RESET test had
been applied using a higher power no less than six, the power curve of
``Triple-U'' shape will have little difference to other shapes. The
recommendation of the polynomial power for the RESET should be revised,
perhaps. This illustrates the sensitivity of conventional testing to the
parameters, and it also points to a limitation that one needs to know
the data generating process in order to set the parameters for the test.

For visual tests, based on the orders of the Hermite polynomials, we
expect the ``U'' shape will be the easiest one to be detected by
subjects followed by the ``S'', ``M'' and ``Triple-U'' shape. From
Figure \ref{fig:poly-power-uniform-j}, it can be observed that the power
curves are mostly aligned with our expectation, except for the ``M''
shape, which is as easy to be detected as the ``U'' shape. This implies
the readability of the shape do not strictly follow the degree of the
polynomials.

@@

We have also investigated the impact of different heteroskedasticity
shapes on power of conventional tests and visual tests. In theory, the
``Left-triangle'' and the ``Right-triangle'' shapes are functionally
identical from the point of view of a Breusch-Pagan test. As shown in
Figure \ref{fig:heter-power-uniform-a}, this is indeed the case where
little difference between the power curves can be perceived. Similarly,
visual tests should have the smae power in detecting these two shapes if
they are equally likely to be identified. However, it can be observed
from Figure \ref{fig:heter-power-uniform-a} that the power curve for the
``Left-triangle'' shape is constantly higher than the one for the
``Right-triangle'' shape, indicating a potential favour of orientation
by human, which is worth to be explored in future studies. Besides, the
chance of detecting the ``Butterfly'' shape by visual tests are slightly
higher than other two shapes.

@@

Figure \ref{fig:different-x-dist-poly-power} illustrates the change of
power of visual tests and conventional tests for different fitted
distributions used in the non-linearity model and the heteroskedasticity
model. Note that we have collected more evaluations on lineup with
uniform fitted value distribution, and the number of evaluations on a
lineup will have an impact on the power of the visual test. To have a
fair comparison, the five evaluations used in the power estimation are
randomly sampled from the total eleven evaluations for each lineup with
uniform fitted value distribution. The sampling process has been
repeated 500 times to visualize the variation.
